---
version: draft
author: "Michael Garrett"
date: "`r Sys.Date()`"
title: "Descriptive statistics for attitude toward STEM data"
output: 
  rmarkdown::html_document:
    css: ../css/simple_report.css
vignette: >
  %\VignetteIndexEntry{attitude-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
## Load libraries

library(tidyverse)  # map, mutate, filter, gather, select, group_by, summarize
library(broom)
library(knitr)
library(algor)

## RMD options

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  fig.width = 7, 
  fig.height = 5,
  fig.path = "../figures/"
)

## Set local vignette parameters

# NA
```

# Get data

Retrieving data object and creating tidy stacked version

```{r get-dat}

## Read dataset from package data into "Global" environment

# Attitude towards STEM in HS. Make composite scores for math, cs, and stem
load(file.path("..", "data", "attitude_ihs_tbl.rda"))
attitude_ihs_tbl <- attitude_ihs_tbl %>% 
  mutate(math = (calc + trig)/2, csci = comp_prog, 
         stem = rowMeans(select(., calc:eng), na.rm = TRUE))

# Attitude towards STEM work after HS. Make composite scores for cs, and stem
load(file.path("..", "data", "attitude_ahs_tbl.rda"))
attitude_ahs_tbl <- attitude_ahs_tbl %>% 
  mutate(csci = (cs + prog)/2, 
         stem = rowMeans(select(., cs:eng), na.rm = TRUE))

```

# Descriptive statistics

```{r descriptives-IN-HS}

## IN HIGH SCHOOL mean/SD for attitudes toward stem courses/subjects 
#
# Summarize scores for courses only (raw data)
ihs_course_desc <- attitude_ihs_tbl %>% 
  select(-c(ID, phys, math, csci, stem)) %>% 
  group_by(timing) %>% 
  summarize_all(list(M = mean, S = sd), na.rm = TRUE) 

# Nicely formatted display table for report
ihs_course_desc %>%
  kable(digits = 2, 
        caption = "Mean attitude scores on IN HS measures by class")

# Only include composites measures
ihs_comp_desc <- attitude_ihs_tbl %>% 
  select(c(timing, phys, math, csci, stem)) %>% 
  group_by(timing) %>% 
  summarize_all(list(M = mean, S = sd), na.rm = TRUE)

# Display table for report
ihs_comp_desc %>%
  kable(digits = 2, 
        caption = "Mean attitude on IN HS composite measures")

```

```{r descriptives-AFTER-HS}

## AFTER HIGH SCHOOL mean/SD for attitudes toward stem study/work
#
# Summarize scores for courses only (raw data)
ahs_area_desc <- attitude_ahs_tbl %>% 
  select(-c(ID, phys, math, csci, stem)) %>% 
  group_by(timing) %>% 
  summarize_all(list(M = mean, S = sd), na.rm = TRUE) 

# Formatted table for report
ahs_area_desc %>%
  kable(digits = 2, 
        caption = "Mean attitude scores on AFTER HS measures by field")

# Only include composites measures
ahs_comp_desc <- attitude_ahs_tbl %>% 
  select(c(timing, phys, math, csci, stem)) %>% 
  group_by(timing) %>% 
  summarize_all(list(M = mean, S = sd), na.rm = TRUE) 

ahs_comp_desc %>%
  kable(digits = 2, 
        caption = "Mean attitude scores on AFTER HS composite measures")
```

# Illustrate change over intervention in key areas

```{r attiude-graph-setup}

# These captions describe the way composite measures were built. They will be
# attached to the figures below.

ihs_cap <- "csci corresponds to 'computer programming', math averages 'calculus and trig', stem averages of calculus, trig, physics, advanced chemistry, and computer programming"

ahs_cap <- "csci averages 'programming' and 'computer science', stem averages 'math', 'physics', 'chem', 'programming' and 'computer science'"

```

## Attiutde towards classes in high school

### Group mean attitudes pre/post

```{r iha-mean-attitude-graph}

## Attitude toward subjects IN HS

# This plot will show pre and post scores for each composite measure as two
# points connected by a line. Subject will be coded as color. A marginal gloss
# defines the composites. Scores are described in a brief caption split to fit
# in a marginal gloss

ihs_comp_desc %>% 
  select(timing, ends_with("_M")) %>% 
  gather(-timing, key = subject, value = score) %>% 
  ggplot(aes(x = timing, y = score, col = subject, group = subject)) +
    geom_line(size = 1.5) + geom_point(size = 3) +
    ylim(1, 4) +
    coord_fixed(ratio = 1.25, xlim = c(1.35, 1.65), clip = "off") +
    labs(title = "Pre/post mean attitude toward taking a subject in highschool",
         subtitle = "attitude from 'absolute no' (1) to 'absolute yes' (2).") + 
    annotate("text", label = paste(strwrap(ahs_cap, width = 25), collapse = "\n"),
             x = 2.4, y = 4.0, hjust = 0, vjust = 1, size = 4)

```

### Individual attiudes pre/post

```{r ihs-indiv-attitude-graph}

## Attitude toward subjecets AFTER HS

# This plot will show pre and post scores for each composite measure as two
# points connected by a line. Subject will be coded as color. A marginal gloss
# defines the composites. Scores are described in a brief caption.
attitude_ihs_tbl %>% 
  select(timing, ID, phys, math, csci, stem) %>% 
  gather(-timing, -ID, key = subject, value = score) %>% 
  ggplot(aes(x = timing, y = score, col = subject, group = subject)) +
    geom_line(size = 1.5) + geom_point(size = 3) +
    ylim(1, 4) +
    coord_fixed(ratio = 1.25, xlim = c(1.35, 1.65), clip = "off") + 
    labs(title = "Pre/post individual attitudes toward taking a subject in highschool",
         subtitle = "attitude scores from 'absolute no' (1) to 'absolute yes' (4).",
         caption = paste(strwrap(ahs_cap, width = 80), collapse = "\n")) +
    facet_wrap(~ID)

```

## Attitudes toward working at a subject after high school

Group mean attitudes pre/post

```{r ahs-mean-attitude-graph}

## Attitude toward subjecets AFTER HS

# This plot will show pre and post scores for each composite measure as two
# points connected by a line. Subject will be coded as color. A marginal gloss
# defines the composites. Scores are described in a brief caption split to fit
# in a marginal gloss

ahs_comp_desc %>% 
  select(timing, ends_with("_M")) %>% 
  gather(-timing, key = subject, value = score) %>% 
  ggplot(aes(x = timing, y = score, col = subject, group = subject)) +
    geom_line(size = 1.5) + geom_point(size = 3) +
    ylim(1, 4) +
    coord_fixed(ratio = 1.25, xlim = c(1.35, 1.65), clip = "off") +
    labs(title = "Pre/post mean attitude toward working at a subject after highschool",
         subtitle = "attitude scores from 'absolute no' (1) to 'absolute yes' (4).") +
    annotate("text", label = paste(strwrap(ahs_cap, width = 25), collapse = "\n"),
             x = 2.4, y = 4.0, hjust = 0, vjust = 1, size = 4)

```

### Individual attiudes pre/post

```{r ahs-indiv-attitude-graph}

## Attitude toward subjecets AFTER HS

# This plot will show pre and post scores for each composite measure as two
# points connected by a line. Subject will be coded as color. A marginal gloss
# defines the composites. Scores are described in a brief caption.
attitude_ahs_tbl %>% 
  select(timing, ID, phys, math, csci, stem) %>% 
  gather(-timing, -ID, key = subject, value = score) %>% 
  ggplot(aes(x = timing, y = score, col = subject, group = subject)) +
    geom_line(size = 1.5) + geom_point(size = 3) +
    ylim(1, 4) +
    coord_fixed(ratio = 1.25, xlim = c(1.35, 1.65), clip = "off") + 
    labs(title = "Pre/post individual attitude working at a subject after highschool",
         subtitle = "attitude scores from 'absolute no' (1) to 'absolute yes' (4).",
         caption = paste(strwrap(ahs_cap, width = 80), collapse = "\n")) +
    facet_wrap(~ID)

```

